Here is the list of all the movies
my algorithm is based on this principle: If an user correctly rate a movie then
he will also rate the movie that a looking like the ones he has already rated.

- Between all the movies, What are the movies that are similar to each others?
- If I would like to predict the rating an user would give on a movie, then, I will 
start by finding all the movies that are similar to that movie (movie to predict ratings)
The ratings of the user will be the mean of the ratings that same user give to the movies similar to 
that movie to predict the ratings

The issue will that approach is that :
  the data give to us is the ratings data. It means that we are supposed to 
  split the ratings into train/test 
  But in the algorithm described above no splitting is done
  
A movie is defined by:
  - its own properties: release date, genre, 
  - a vector of rating from users. 


```{r message=FALSE, warning=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")

if(!require(cmfrec)) install.packages("cmfrec", repos = "http://cran.us.r-project.org")
if(!require(logger)) install.packages("logger")
if(!require(stringr)) install.packages("stringr", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2")

library(tidyverse)
library(caret)

library(cmfrec)
library(logger)
library(stringr)
library(ggplot2)
```

# Create train and final holdout test sets

```{r}
source("create_train_and_final_holdout_test_sets.R")
```

# Multiple hot encoding of Genre

```{r}
log_info("Apply multiple hot encoding on movie Genre")

movie_genres <- c("Action", "Adventure", "Animation", "Children", "Comedy",
                  "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir",
                  "Horror", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller",
                  "War", "Western")

genres <- edx$genres

split_genre <- function(genre) {
  s <- unlist(str_split(genre, fixed("|")))
  ix <- rep(0, length(movie_genres))
  ix[match(s, movie_genres)] <- 1
  
  return(ix)
}

genres.splitted <- do.call(rbind, lapply(genres, FUN=split_genre))
colnames(genres.splitted) <- movie_genres
```

```{r}
log_info("Multiple hot encoding applied on movie [genre]")
log_info("Extract [year] & [title] from movie [name]")

movies.data <- genres.splitted
movies.name.splitted <- edx %>%
  mutate(
    year = as.integer(str_extract(title, "(?<=\\()\\d+(?=\\)$)")),
    name = str_squish(str_remove(title, " \\(\\d+\\)$"))
  ) %>% select(name, year)

movies.data <- cbind(movies.data, year = movies.name.splitted$year)

log_info("[year] & [name] extracted from movie name")

edx.data <- cbind(edx, movies.data) %>%
  select(-c(title, genres))

log_info("Added new features (Multiple-hot encoding Genre) to EDX data.frame")
log_info("Selected all features from EDX data.frame except [title] and [genres] features")
```


```{r}
remove(movies.data, movies.name.splitted, genres, genres.splitted)
```

# Training

## Create Train and Test sets

```{r}
log_info("Split the EDX dataset to 80% of training and 20% of testing")

set.seed(1) # if using R 3.5 or earlier
test_index <- createDataPartition(y = edx.data$rating, times = 1, p = 0.25, list = FALSE)
train.edx <- edx.data[-test_index,]
temp <- edx.data[test_index,]

log_info("Make sure userId and movieId in final hold-out test set are also in edx set")

test.edx <- temp %>%
  semi_join(train.edx, by = "movieId") %>%
  semi_join(train.edx, by = "userId")

log_info("Add rows removed from final hold-out test set back into edx set")

removed <- anti_join(temp, test.edx)
train.edx <- rbind(train.edx, removed)

rm(test_index, temp, removed)
```

## Prepare data set and loss function

```{r}
log_info("Prepare data set and loss function")

rmse <- function(y, y_) {
  return(sqrt(mean((y - y_)^2)))
}

X.train <- train.edx %>% select(userId, movieId, rating)
X.test <- test.edx %>% select(userId, movieId, rating)

X.movies.data <- train.edx %>% select(-c(userId, movieId, rating, -timestamp))

```

## Continue Without movies data

```{r}
log_info("MATRIX FACTORIZATION Without movies data")
```

### Finding the best number of factors k

```{r}
log_info("Finding the best number of factors k")

errors <- c()
best_model <- NULL
best_error <- 10e6

startTime <- Sys.time()
for (k in 3:50) {
  log_info("Start k={k}")
  model <- CMF(X.train, k = k, method = 'lbfgs',
               user_bias = TRUE, item_bias = TRUE,
               center = TRUE,
               # NA_as_zero = TRUE, 
               nthreads = 1, verbose = FALSE, seed = 1)
  
  predictions <- predict(model, user=X.test$userId, item=X.test$movieId)
  k_errors <- rmse(X.test$rating, predictions)
  
  if (k_errors < best_error) {
    best_model <- model
    best_error <- k_errors
  }
  
  errors <- c(errors, k_errors)
  log_info("End k={k}; Error={k_errors}")
}
endTime <- Sys.time()

print(paste0("Duration: ", (endTime - startTime)))
```

### Plot the errors

Let's plot the errors to see the one with the minimum RMSE

```{r}
errors.df <- data.frame(k = 3:50, error = errors)
ggplot(data=errors.df, aes(x = k, y = error)) + geom_line() + geom_point()

k.min <- which.min(errors)
print(paste0("k.min: ", k.min))

remove(errors.df, best_error, best_model, errors)
```

### Train the all dataset

Now that we have the best value for `k`, we have to train the whole dataset with the number of factor that give us the best RMSE in the previous step. The new model will be trained using the whole `edx` dataset.

```{r}
model <- CMF(X, k = k.min, method = 'lbfgs',
               user_bias = TRUE, item_bias = TRUE,
               center = TRUE,
               # NA_as_zero = TRUE, 
               nthreads = 1, verbose = FALSE, seed = 1)
```

### Predictions on final holdout dataset

```{r}
X.fht <- final_holdout_test %>% select(userId, movieId, rating)
predictions <- predict(model, user=X.fht$userId, item=X.fht$movieId)
fht_error <- RMSE(X.fht$rating, predictions)
print(paste0("The predictions error is", fht_error))
```

The predictions error is `r fht_error`


## With movies data

```{r}
log_info("MATRIX FACTORIZATION WITH movies data")
```

### Finding best number of factors (k)

To find the best number of factor, for each value of k, we will train the model with `X.train`, then use the model obtained to predict value from `X.test` and keep the prediction RMSE.
The best number of factor will be the one with the minimum RMSE.


```{r}

log_info("Finding the best number of factors k")

errors <- c()
best_model <- NULL
best_error <- 10e6

startTime <- Sys.time()
for (k in 3:30) {
  log_info("Start k={k}")
  model <- CMF(X.train, k = k, I=X.movies.data, method = 'lbfgs',
               user_bias = TRUE, item_bias = TRUE,
               center = TRUE,
               # NA_as_zero = TRUE, 
               nthreads = 1, verbose = FALSE, seed = 1)
  
  predictions <- predict(model, user=X.test$userId, item=X.test$movieId)
  k_errors <- rmse(X.test$rating, predictions)
  
  if (k_errors < best_error) {
    best_model <- model
    best_error <- k_errors
  }
  
  errors <- c(errors, k_errors)
  log_info("End k={k}; Error={k_errors}")
}
endTime <- Sys.time()

print(paste0("Duration: ", (endTime - startTime)))
```

### Plot the errors

Let's plot the errors to see the one with the minimum RMSE

```{r}
errors.df <- data.frame(k = 3:50, error = errors)
ggplot(data=errors.df, aes(x = k, y = error)) + geom_line() + geom_point()

k.min <- which.min(errors)
print(paste0("k.min: ", k.min))

remove(errors.df, best_error, best_model, errors)
```


### Train the all dataset

Now that we have the best value for `k`, we have to train the whole dataset with the number of factor that give us the best RMSE in the previous step. The new model will be trained using the whole `edx` dataset.

```{r}
X <- edx.data %>% select(userId, movieId, rating)
I <- edx.data %>% select(-c(userId, movieId, ratings_file))

model <- CMF(X, I = I, k = k.min, method = 'lbfgs',
               user_bias = TRUE, item_bias = TRUE,
               center = TRUE,
               # NA_as_zero = TRUE, 
               nthreads = 1, verbose = FALSE, seed = 1)
remove(X)
```

### Predictions on final holdout dataset

```{r}
X.fht <- final_holdout_test %>% select(userId, movieId, rating)
predictions <- predict(model, user=X.fht$userId, item=X.fht$movieId)
fht_error <- rmse(X.fht$rating, predictions)
print(paste0("The predictions error is", fht_error))
```

The predictions error is `r fht_error`

```{r}
log_info("END")
```

